The complex network of trophic interactions in a subAntarctic Marine Protected

### 2023-01-30 MS_ToSubmit/MS_Burdwood_foodweb.Rmd
Finish Intro

keywords: 
  - Food web
  - Complexity
  - Structure
  - Marine Protected Area
  - Southwest Atlantic
journal: "MEPS, Oikos, Polar Biology (?)"
date: "`r Sys.Date()`"
classoption: preprint, 3p, authoryear
bibliography: AMPBancoBurdwood.bib
nocite: '@*'
linenumbers: true
numbersections: true
# Use a CSL with `citation_package = "default"`
csl: "apa-old-doi-prefix.csl"
output: 
  rticles::elsevier_article:
    keep_tex: true
    citation_package: "default"
---

@@

* Introduction to MPAs and MPA N-BB

@@

* Ecological importance of MPA N-BB to the region

Many of these MPAs centre on the presence of particularly vulnerable, keystone, or charismatic species; the presence of large numbers (or proportions) of endemic species; and the presence of high biodiversity across taxonomic levels [@Hogg2016]. Indeed, the MPA N-BB has been created because it is considered a biodiversity hotspot, mostly of its benthic community [@Schejter2016; @Falabella2017]. It is characterized by a high biomass of vulnerable and fragile species (mainly Porifera and Cnidaria) that meet the characteristics of vulnerable marine ecosystems [@Schejter2021], here defined as sites that present taxa with a minimum biomass of 10 kg per 1200 m^2^ [@CommissionfortheConservationofAntarcticMarineLivingResources2009]. Also, the benthic community provides habitat to several small-sized species [@Lopez-Gappa2018; @MartinSirito2019; @Schejter2019], and has an important role in the life history of fishes as a food source, refuge, or nursery area [@GarciaAlonso2018; @Delpiani2020; @Troccoli2020]. The maintenance of this singular community is related to the oceanographic processes, including the circulation of the rich Malvinas (Falkland) current in the area [@Piola1989; @Guerrero1999] and the upwelling and mixing phenomena [@Matano2019]. The input of nutrients coming from the Malvinas (Falkland) current also supports a diverse plankton community [@Guinder2020]. Overall, 811 benthic and planktonic species have been identified for the MPA N-BB ecosystem, where 349 were reported for the first time for the MPA [@APN2022]. At the ecosystem level the MPA N-BB is ascribed as a carbon-sinker [@Bergagna2019].

* Trophic ecology & keystone spp knowledge of B Burdwood

The structure and functioning of ecosystems can be successfully described by studying one of the most-frequent interactions occurring between species: the predator-prey relationship [@Bascompte2009]. The predator-prey or trophic interactions of a particular region are commonly referred as the food web, that represents the roadmap for matter and energy flow in an ecosystem.
The trophic structure of the southwestern South Atlantic Ocean has been proposed to be under a 'wasp-waist' control, meaning that the structure and dynamics of the ecosystem is regulated primarily by mid-trophic level species (e.g. fishes, crustaceans) [@Padovani2012; @Saporiti2015; @Riccialdelli2020]. In particular, the ecosystem of the MPA N-BB shows a more pronounced 'wasp-waist' structure than other subAntarctic areas, like the Atlantic coast of Tierra del Fuego, where the Fuegian sprat *Sprattus fuegensis* and longtail southern cod *Patagonotothen ramsayi* are considered the most plausible 'wasp-waist' species [@Riccialdelli2020].
Recently, the identification of the most influential species within food webs has become a crucial task for conservation purposes in local and regional contexts to maintain ecosystem integrity and the supply of ecosystem services for the southwestern South Atlantic Ocean [@Franco2020; @Bergagna2020; @Riccialdelli2020].

* Why using a network approach

In the past decades, several researches have shown that high-latitude marine ecosystems, such as the MPA N-BB, are characterised by an unexpected complexity in terms of biodiversity and ecological interactions [@Day2013; @Kortsch2019; @Cordone2020; @Trathan2021].

Anthropogenic and climate-induced impacts are known to affect individual species, but they can also change the interactions that each species has with its prey and predators (Constable et al. 2014, Lynam et al. 2017).

* Objective

In the present work we present the first, detailed analysis of the food web for the MPA N-BB ecosystem. For this we applied a network approach to a highly-resolved food web. The objective was twofold: characterise the food web in terms of complexity and structure, and describe the species' role in such network framework.

@@

We found different relationships between the species trophic level (TL) and the rest of the analysed species-level properties (Figure 4A-D). In this regard, the most evident significant relationship was with trophic similarity, i.e. the higher the species' TL, the lower the trophic similarity or the higher the uniqueness in terms of trophic role (Figure 4C). Here it's noteworthy to highlight high-trophic level species with low values of trophic similarity: *Bathyraja macloviana* and *Squalus acanthias* (Chondrichthyans), *Diplopteraster clarki* and *Pteraster sp* (echinoderms), *Phalacrocorax_atriceps* and *Eudyptes chrysocome* (sea birds), and *Lagenorhynchus cruciger* and *Mesoplodon bowdoini* (marine mammals) (Table S3).

@@

It's noteworthy that the highest values of betweenness were shown by species of mid-TLs (3-4), meaning that those species participated in the highest number of shortest paths between species (Figure 4A). These were the species with the highest values: *Patagonotothen ramsayi*, *Dissostichus eleginoides*, *Salilota australis* (fishes), *Doryteuthis gahi* (Cephalopoda), and *Patagonotothen guntheri* (Notothenioid fish) (Table S3).

@@

This study was funded by Consejo Nacional de Investigaciones Científicas y Técnicas (CONICET) and Agencia Nacional de Promoción Científica y Tecnológica (PICT 2020-SERIEA-01617). This work is contribution no. XX of the MPA Namuncurá (Law 26875).

>>>>

keywords: 
  - Food web
  - Complexity
  - Structure
  - Marine Protected Area
  - Southwest Atlantic
journal: "MEPS, Oikos, Ecological Applications, J of Applied Ecology (?)"
date: "`r Sys.Date()`"
classoption: preprint, 3p, authoryear
bibliography: AMPBancoBurdwood.bib
nocite: '@*'
linenumbers: true
numbersections: true
# Use a CSL with `citation_package = "default"`
csl: "apa-old-doi-prefix.csl"
output: 
  rticles::elsevier_article:
    keep_tex: true
    citation_package: "default"
---

@@

@@

Many of these MPAs centre on the presence of particularly vulnerable, keystone, or charismatic species; the presence of large numbers (or proportions) of endemic species; and the presence of high biodiversity across taxonomic levels [@Hogg2016]. Indeed, the MPA N-BB was created because it is considered a biodiversity hotspot, mostly of its benthic community [@Schejter2016; @Falabella2017]. This community is featured by a high biomass of vulnerable and fragile species (mainly Porifera and Cnidaria) that meet the characteristics of vulnerable marine ecosystems [@Schejter2021], here defined as sites that present taxa with a minimum biomass of 10 kg per 1200 m^2^ [@CommissionfortheConservationofAntarcticMarineLivingResources2009]. Also, it provides habitat to several small-sized species [@Lopez-Gappa2018; @MartinSirito2019; @Schejter2019], and has an important role in the life history of fishes as a food source, refuge, or nursery area [@GarciaAlonso2018; @Delpiani2020; @Troccoli2020]. The maintenance of this singular community is related to the oceanographic processes, including the circulation of the rich Malvinas (Falkland) current in the area [@Piola1989; @Guerrero1999] and the upwelling and mixing phenomena [@Matano2019]. The input of nutrients coming from the Malvinas (Falkland) current also supports a diverse plankton community [@Guinder2020]. Overall, 811 benthic and plankton species have been identified for the MPA N-BB ecosystem, where 349 were reported for the first time for the area [@APN2022].

Recently, the identification of the most influential species has become a crucial task for conservation purposes in local and regional contexts to maintain ecosystem integrity and the supply of ecosystem services for the southwestern South Atlantic Ocean [@Franco2020; @Bergagna2020; @Riccialdelli2020].
In this regard, the structure of the southwestern South Atlantic Ocean has been proposed to be under a 'wasp-waist' control, meaning that the structure and dynamics of the ecosystem is regulated primarily by mid-trophic level species (e.g. fishes, crustaceans) [@Padovani2012; @Saporiti2015; @Riccialdelli2020]. In particular, the ecosystem of the MPA N-BB shows a more pronounced 'wasp-waist' structure than other subAntarctic areas, like the Atlantic coast of Tierra del Fuego, where the Fuegian sprat *Sprattus fuegensis* and longtail southern cod *Patagonotothen ramsayi* are considered the most plausible 'wasp-waist' species [@Riccialdelli2020].

High-latitude marine ecosystems, such as the MPA N-BB, are characterised by an unexpected complexity in terms of biodiversity and ecological interactions [@Day2013; @Kortsch2019; @Cordone2020; @Trathan2021]. Although there is a robust knowledge about the complexity considering the richness of the benthic and plankton communities in the MPA N-BB ecosystem [e.g. @Schejter2016; @Guinder2020; @APN2022], there is a need for a better understanding of species interactions' complexity and structure.
This aspect can be tackled by analysing one of the most-frequent relationships occurring between species: the predator-prey interaction [@Bascompte2009]. The sum of predator-prey or trophic interactions of a particular region is referred to as a food web, which represents the roadmap for matter and energy flow in an ecosystem.
In recent years, network approaches have been successfully applied to study complex high-latitude marine ecosystems, improving our knowledge on structure, functioning, and respond to environmental/anthropogenic changes [e.g. @Kortsch2015; @Cordone2018; @Funes2022; @Marina2023]. Here it is worth mentioning that anthropogenic threats have been recently reported for the MPA N-BB region [@Cossi2021; @Fioramonti2022].

In the present work we present the first, detailed analysis of the network of predator-prey interactions, hereafter food web, for the MPA N-BB ecosystem. For this we applied a network approach to a highly-resolved food web. The objective was twofold: characterise the food web in terms of complexity and structure, and identify the species' role in such network framework.

@@

We found different relationships between the species trophic level (TL) and the rest of the analysed species-level properties (Figure 4A-D). In this regard, the most evident significant relationship was with trophic similarity, i.e. the higher the species' TL, the lower the trophic similarity or the higher the uniqueness in terms of trophic role (Figure 4C). Here it is noteworthy to highlight high-trophic level species with low values of trophic similarity: *Bathyraja macloviana* and *Squalus acanthias* (Chondrichthyans), *Diplopteraster clarki* and *Pteraster sp* (echinoderms), *Phalacrocorax_atriceps* and *Eudyptes chrysocome* (sea birds), and *Lagenorhynchus cruciger* and *Mesoplodon bowdoini* (marine mammals) (Table S3).

@@

It is noteworthy that the highest values of betweenness were shown by species of mid-TLs (3-4), meaning that those species participated in the highest number of shortest paths between species (Figure 4A). These were the species with the highest values: *Patagonotothen ramsayi*, *Dissostichus eleginoides*, *Salilota australis* (fishes), *Doryteuthis gahi* (Cephalopoda), and *Patagonotothen guntheri* (Notothenioid fish) (Table S3).

@@

This study was funded by Consejo Nacional de Investigaciones Científicas y Técnicas (CONICET) and Agencia Nacional de Promoción Científica y Tecnológica (PICT 2020-SERIEA-01617), Argentina. This work is contribution no. XX of the MPA Namuncurá (Law 26875).
